Architectures for workflow integration in science gateways

### 2016-04-13 paper.tex
Updates in section 3.1

Figure~\ref{fig:notations} shows the graphical notations used to
describe the architectures in the reminder of this
Section. Architectures are described from their main software
components and interactions.

\todo{add interaction a with the infrastructure for all architectures (application porting)}

@@

The interactions described below are labeled consistently with the
notations used on Figures~\ref{archi:tight}-\ref{archi:import}.

@@

\item Workflow integration: consists in adding a new workflow to the
  science gateway so that users can execute it. It is conducted by
  science gateway administrators, developers or users, and it results
  in an interface, for instance a web form, where users can enter the
  parameters of the workflow to be executed. Integrating a workflow is
  not the same process as integrating a workflow engine.
\item Task control: operations to manage tasks on the infrastructure,
  including: submission, monitoring, termination, deletion, etc.
\item Data control: operations to manage data on
  the infrastructure, such as: upload, download, deletion, browsing,
  replication, caching, etc. Data movements can be triggered by the
  user, to upload input data or download processed data -- interaction
  \texttt{$c_1$}, or by the workflow engine, to transfer data across the
  infrastructure -- interaction \texttt{$c_2$}. \todo{add c1 and c2 to the
    Figures}
\item Workflow control: operations to execute a workflow, including:
  submission, monitoring, termination, etc. Workflow control can be
  coarse-grained (a.k.a. black box) or fine-grained (white box). In a
  coarse-grained model, workflow activities are masked. 
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc.
\item Pool-agent: specific to the architecture described in~\ref{sec:pool}.
\item Workflow conversion: translation from one workflow language to
  another one.
\end{enumerate}

>>>>

Figure~\ref{fig:notations} shows the graphical notations used to
describe the architectures in the reminder of this
Section. Architectures will be described from their main software
components and interactions. Software components include science
gateway and infrastructure in all architectures, and workflow service,
workflow pool and agent are involved in some architectures. The
workflow engine itself is represented by a specific symbol. Software
interactions among these components are described in the next
Section. Abstract interactions are specific types of interactions that
may be implemented by various different software interactions. Red
color is used to represent the components and interactions that need
to be specifically developed to integrate workflow engines. Tasks
refer to computing tasks that are created by the workflow engine and
executed on the infrastructure. Data represents any type of file, or
database that is involved in the workflow execution and stored on the
infrastructure while the workflow executes. It does not cover workflow
parameters such as strings, numbers, etc.

@@

The interactions involved in the architectures are described below and
labeled consistently with the notations used on
Figure~\ref{fig:architectures}.

@@

\item Workflow integration: consists in adding a new workflow to the
  system so that users can execute it. It is triggered by science
  administrator and it results in an interface, for instance a web
  form, where users can enter the parameters of the workflow to be
  executed. The interaction has two aspects: (\texttt{a$_1$}) the
  programs used in the workflow are installed on the infrastructure,
  which may or may not require administrative privileges on the
  infrastructure; (\texttt{a$_2$}) the workflow has to be configured
  in the science gateway so that it becomes available to users. Note
  that integrating a workflow is not the same process as integrating a
  workflow \emph{engine}.
\item Task control: operations to manage tasks on the
  infrastructure, including: authentication, submission, monitoring,
  termination, deletion, etc. When the infrastructure is a grid or a
  cloud, it is for instance implemented using libraries such as SAGA,
  DRMAA, OCCI and similar initiatives \todo{refs needed}.
\item Data control: operations to manage data on the infrastructure,
  such as: upload, download, deletion, browsing, replication, caching,
  etc. Data movements can be triggered by the user in the science
  gateway (\texttt{$c_1$}), to upload input data or download processed
  data. They can also be performed by the workflow engine
  (\texttt{$c_2$}), to transfer data across the
  infrastructure. 
\item Workflow control: operations to execute a workflow with an
  engine, including: workflow submission, monitoring, termination,
  etc. Workflow control can be coarse-grained (a.k.a. black box) or
  fine-grained (white box). In a coarse-grained model, workflow
  activities are masked and the user only has a global view of the
  workflow execution. In a fine-grained model, user is exposed to the
  workflow topology, i.e. to the outputs of the individual tasks,
  their statuses and so on. 
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc. Sub-task control is similar to
  interaction \texttt{b}, except that information about the parent
  task which submitted a sub-task is usually available and used for
  additional control. For instance, the parent task may wait for all
  its sub-tasks to finish before finishing, and conversely all the
  sub-tasks may be killed if the parent task is killed.
\item Pool-agent: specific to the architecture described
  in Section~\ref{sec:pool}. This is an interaction used when agents retrieve
  work from a central pool. It covers agent registration and
  de-registration to the pool, protocols to send work from the pool to
  the agent, mechanisms to update work status, and so on. 
\item Workflow conversion: translation from one workflow language to
  another one. This interaction may not be available or implementable
  for every workflow language. It has been developed mostly for
  well-structured and relatively simple workflow language such as between GWorkflowDL
  and Scufl~\cite{OLAB-09} and between the 4 languages that were involved in the
  SHIWA initiative. In SHIWA, workflow conversion is done through an
  intermediate representation, the Interoperable Workflow Intermediate
  Representation~\cite{plankensteiner-montagnat-etal:2011}, which
  allows to convert among $n$ workflow languages using $2n$
  interactions instead of $n^2$.
\end{enumerate}
